---
title: "1. Biodiversity effects over time"
author: "meganedeziel"
date: "2023-09-27"
output:
  pdf_document: default
  html_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, warning=FALSE, message=FALSE)
```

### package loading
```{r, messages = FALSE}
library(here)
library(tidyverse)
library(nlme)
library(emmeans)
library(ggpubr)
library(grid)
library(car)
library(rstatix)
library(formatR)
```

### data importation
```{r}
#Data on Net effect, CE, SE 
NE_data<-read.csv(here("data", "NE_data.csv"), sep=";")

#transform Composition, Block, Days as factor for further analyses
NE_data<-NE_data[NE_data$Days!="0", -1]
NE_data<-NE_data %>%
  mutate(across(c("Composition", "Block", "Days"), 
                as.factor)) 
str(NE_data)
```

### test homoscedasticity of variance
```{r}
# Plot Raw data - NE against Days
PlotNE <- ggplot(NE_data, aes(x = Days, y = NE)) + 
  geom_jitter(position=position_jitter(0.2)) + 
  stat_summary(fun.data="mean_sdl",  fun.args = list(mult=1), 
               geom="pointrange", color = "red")+
  theme_classic() +
  labs(x="Days", y="NE (mg)") +
  scale_y_continuous(limits=c(-2000, 2000), breaks=seq(-2000, 2000, by=500)) 

leveneTest(NE ~ Days, data=NE_data) #the variance among the factor Days is NOT equal

PlotCE <- ggplot(NE_data, aes(x = Days, y = CE)) + 
  geom_jitter(position=position_jitter(0.2)) + 
  stat_summary(fun.data="mean_sdl",  fun.args = list(mult=1), 
               geom="pointrange", color = "red")+
  theme_classic() +
  labs(x="Days", y="CE (mg)") +
  scale_y_continuous(limits=c(-2000, 2000), breaks=seq(-2000, 2000, by=500))

leveneTest(CE ~ Days, data=NE_data) #variance among the factor Days is NOT equal

PlotSE <- ggplot(NE_data, aes(x = Days, y = SE)) + 
  geom_jitter(position=position_jitter(0.2)) + 
  stat_summary(fun.data="mean_sdl",  fun.args = list(mult=1), 
               geom="pointrange", color = "red")+
  theme_classic() +
  labs(x="Days", y="SE (mg)") +
  scale_y_continuous(limits=c(-2000, 2000), breaks=seq(-2000, 2000, by=500))

leveneTest(SE ~ Days, data=NE_data) #variance among the factor Days is NOT equal

#mean_sdl computes the mean plus or minus a constant times the standard deviation 
#(here the constant (mult) =1), so we see mean + standard deviation in Red
#We can clearly see that the variance of NE, CE and SE increases over time 

ggarrange(PlotNE, PlotCE, PlotSE, nrow=1, ncol=3)
```


### why we use the nlme::lme function to compute the ANOVAs
Since the variance of NE, CE, SE increases over time (heteroscedasticity in the residuals), we decided to use the function nlme::lme to make our ANOVAs, because the function allows to weight the variance by the factor Days. Observations with higher variability will have lower weights, allowing the model to give them less influence. This method is particularly relevant with repeated measures data as we have here. 

The class pdMat, which represent positive-definite matrices, is used to represent variance-covariance matrices of random effects. 
We have have two blocks in the random effects variance-covariance matrix: one for Composition, and one for Block.

Sources: https://stackoverflow.com/questions/36643713/how-to-specify-different-random-effects-in-nlme-vs-lme4
https://biostatmatt.com/archives/2718

The varIdent function allows different variances according to the levels of a classification factor. 

### ANOVA 1 : net effect through time
```{r}
# Model testing the effect of Days on NE with Composition and Block 
# as random factors and weighted by Days.

# Here we add a dummy factor just to include the whole dataset in a single block,
# to act as a grouping factor within which Composition and Block are both nested. 
NE_data$dummy<-factor(1)

#1.1-ASSESSING WHETHER THERE IS ANY STATISTICALLY SIGNIFICANT DIFFERENCE AMONG DAYS
# this first p-value aims to understand whether there is any statistically significant 
# difference among a set of treatments (Days). In the model, pdBlocked combines the 
# covariance structures of pdIdent Composition and pdIdent Block (identity matrices) 
# in a 2-block matrix. -1 means that there is no default overall intercept being 
# estimated for the factors Composition and Block independently. pdIdent indicates that 
# the random effects are assumed to have equal variances and no covariances.This whole
# syntax is actually a "trick" to specify non-nested crossed random effects in the nlme
# syntax.An equivalent way to specify the random effects could be : 
# random=list(dummy=pdBlocked(list(pdIdent(~ 0 + Composition), pdIdent(~ 0 + Block)))),
modNE <- lme(NE ~ Days,
              random=list(dummy=pdBlocked(list(pdIdent(~Composition-1), 
                                               pdIdent(~Block-1)))),
              data=NE_data,
              weights=varIdent(form=~ 1 | Days), #residuals are allowed to have different 
             #variances for different levels of the "Days" variable
              method="REML", 
              control=list(msMaxIter=1000,
                             msMaxEval=1000))


anova(modNE) #there is at least one statistically significant difference among certain
#levels of factor days 


#1.2-ASSESSING OVERALL EFFECT OF DAYS : likelihood ratio test
# "when assessing the overall treatment effects using a likelihood ratio test, one should 
# use maximum likelihood, rather than REML, when using lme or lmer"
modNE.ml <- lme(NE ~ Days,
              random=list(dummy=pdBlocked(list(pdIdent(~Composition-1), 
                                               pdIdent(~Block-1)))),
              data=NE_data,
              weights=varIdent(form=~ 1 | Days),
              method="ML", 
              control=list(msMaxIter=1000,
                             msMaxEval=1000))
#null model
modNE0.ml <- lme(NE ~ 1,
              random=list(dummy=pdBlocked(list(pdIdent(~Composition-1), 
                                               pdIdent(~Block-1)))),
              data=NE_data,
              weights=varIdent(form=~ 1 | Days),
              method="ML", 
              control=list(msMaxIter=1000,
                             msMaxEval=1000))

#overall p-value modNE
anova(modNE0.ml, modNE.ml) #overall, the factor days has a statistically 
#significant effect on net effect

#1.3-COMPARISON BETWEEN EACH DAY AND DAY 20 (CONTROL LEVEL)
summary(modNE)
#here we can see that NE is significantly greater at day 40 and at day 60 than at day 20

#WHICH TREATMENTS ARE DIFFERENT FROM EACH OTHER? 
contrast(emmeans(modNE, specs="Days"), "pairwise") #the effect of days on NE isn't 
# statistically different between Day 40 and Day 60 only. All other comparisons are significant. 

#MODEL VALIDATION
par(mfrow=c(1,2))
plot(fitted(modNE), resid(modNE,type="pearson"))
qqnorm(resid(modNE,type="pearson"))
qqline(resid(modNE,type="pearson"))

##PLOT
hcl.colors(4, palette="Blue-Yellow")

#computing mean and standard deviation per day from raw data
mean<-as.vector(c(mean(NE_data[NE_data$Days==20, ]$NE), mean(NE_data[NE_data$Days==40, ]$NE), mean(NE_data[NE_data$Days==60, ]$NE)))

se<-as.vector(c(sd(NE_data[NE_data$Days==20, ]$NE)/sqrt(length(NE_data[NE_data$Days==20, ]$NE)), sd(NE_data[NE_data$Days==40, ]$NE)/sqrt(length(NE_data[NE_data$Days==40, ]$NE)), sd(NE_data[NE_data$Days==60, ]$NE)/sqrt(length(NE_data[NE_data$Days==60, ]$NE))))
                                                                                                                                  
NE_meanse<-data.frame(
  days=c("20", "40", "60"),
  mean=mean,
  se=se,
  label=c("a", "b", "b"))
NE_meanse$ci<-NE_meanse$se*1.96
NE_meanse$stars<-NE_meanse$mean + NE_meanse$ci + 125

NEp<-ggplot(NE_meanse) +
  geom_bar(aes(x=days, y=mean), colour="black", fill="#2D3184", stat="identity", alpha=0.7) +
  #geom_errorbar(aes(x=days, ymin=mean-se, ymax=mean+se), width=0.4, colour="red", alpha=0.9, size=1.3) +
  geom_errorbar(aes(x=days, ymin=mean-ci, ymax=mean+ci), width=0.4, colour="black", alpha=0.9, size=1.3) +
  geom_text(aes(x=days, y=mean+ci+100, label=label,), position=position_dodge(0.9), vjust=1, size=7) +
  theme_classic() + 
  theme(text=element_text(size=15), plot.title=element_text(vjust=2), panel.grid.major.y = element_line( size=.1, color="black" )) +
  labs(x="Days of growth", y="NE", title="Net biodiversity effect", pch=8) +
  scale_y_continuous(limits=c(0, 1000), breaks=seq(0, 1000, by=200)) + # Add stars based 
  #on significance levels
  annotate(
    "text",
    x = c(1, 2, 3),  # x-coordinates where stars should be placed
    y = c(320.6969, 603.0179, 750.5188),  # y-coordinate (adjust based on your plot)
    label = c("****", "****", "***"),  # significance labels
    size = 5,  # adjust the size of the stars
    color = "black"  # color of the stars
  )
```

### ANOVA 2 : complementarity effect through time
```{r}
modCE <- lme(CE ~ Days,
              random=list(dummy=pdBlocked(list(pdIdent(~Composition-1), 
                                               pdIdent(~Block-1)))),
              data=NE_data,
              weights=varIdent(form=~ 1 | Days),
              method="REML", 
              control=list(msMaxIter=1000,
                             msMaxEval=1000))
anova(modCE) #there is at least one statistically significant difference among certain 
#levels of factor days 

modCE.ml <- lme(CE ~ Days,
              random=list(dummy=pdBlocked(list(pdIdent(~Composition-1), 
                                               pdIdent(~Block-1)))),
              data=NE_data,
              weights=varIdent(form=~ 1 | Days),
              method="ML", 
              control=list(msMaxIter=1000,
                             msMaxEval=1000))
#null model
modCE0.ml <- lme(CE ~ 1,
              random=list(dummy=pdBlocked(list(pdIdent(~Composition-1), 
                                               pdIdent(~Block-1)))),
              data=NE_data,
              weights=varIdent(form=~ 1 | Days),
              method="ML", 
              control=list(msMaxIter=1000,
                             msMaxEval=1000))

anova(modCE0.ml, modCE.ml) #overall, the factor days has a statistically significant 
#effect on complementarity effect

summary(modCE)
#here we can see that CE is significantly greater at day 40 and at day 60 than at day 20

contrast(emmeans(modCE, specs="Days"), "pairwise") #the effect of days on CE isn't 
#statistically different between Day 20 and Day 40, and between Day 40 and Day 60. 
#All other comparisons are significant. 

par(mfrow=c(1,2))
plot(fitted(modCE), resid(modCE,type="pearson"))
qqnorm(resid(modCE,type="pearson"))
qqline(resid(modCE,type="pearson"))

meanCE<-as.vector(c(mean(NE_data[NE_data$Days==20, ]$CE), mean(NE_data[NE_data$Days==40, ]$CE), mean(NE_data[NE_data$Days==60, ]$CE)))

seCE<-as.vector(c(sd(NE_data[NE_data$Days==20, ]$CE)/sqrt(length(NE_data[NE_data$Days==20, ]$CE)), sd(NE_data[NE_data$Days==40, ]$CE)/sqrt(length(NE_data[NE_data$Days==40, ]$CE)), sd(NE_data[NE_data$Days==60, ]$CE)/sqrt(length(NE_data[NE_data$Days==60, ]$CE))))
                                                                                                                                  
CE_meanse<-data.frame(
  days=c("20", "40", "60"),
  mean=meanCE,
  se=seCE,
  label=c("a", "ab", "b"))
CE_meanse$ci<-CE_meanse$se*1.96
CE_meanse$stars<-CE_meanse$mean + CE_meanse$ci + 125

CEp<-ggplot(CE_meanse) +
  geom_bar(aes(x=days, y=mean), colour="black", fill="#2D3184", stat="identity", alpha=0.7) +
  #geom_errorbar(aes(x=days, ymin=mean-se, ymax=mean+se), width=0.4, colour="red", alpha=0.9, size=1.3) +
  geom_errorbar(aes(x=days, ymin=mean-ci, ymax=mean+ci), width=0.4, colour="black", alpha=0.9, size=1.3) +
  geom_text(aes(x=days, y=mean+ci+100, label=label,), position=position_dodge(0.9), vjust=1, size=7) +
  theme_classic() + 
  theme(text=element_text(size=15), plot.title=element_text(vjust=2), panel.grid.major.y = element_line( size=.1, color="black" )) +
  labs(x="Days of growth", y="CE", title="Complementarity effect", pch=8) +
  scale_y_continuous(limits=c(0, 1000), breaks=seq(0, 1000, by=200)) + 
  annotate(
    "text",
    x = c(1, 2, 3),  
    y = c(296.5644, 433.0912, 958.4401),  
    label = c("****", "****", "***"),  
    size = 5,  
    color = "black"  
  )
```

### ANOVA 3 : Selection effect through time
```{r}
modSE <- lme(SE ~ Days,
              random=list(dummy=pdBlocked(list(pdIdent(~Composition-1), 
                                               pdIdent(~Block-1)))),
              data=NE_data,
              weights=varIdent(form=~ 1 | Days),
              method="REML", 
              control=list(msMaxIter=1000,
                             msMaxEval=1000))
anova(modSE) #there is at least one statistically significant difference among certain
#levels of factor days 

modSE.ml <- lme(SE ~ Days,
              random=list(dummy=pdBlocked(list(pdIdent(~Composition-1), 
                                               pdIdent(~Block-1)))),
              data=NE_data,
              weights=varIdent(form=~ 1 | Days),
              method="ML", 
              control=list(msMaxIter=1000,
                             msMaxEval=1000))
#null model 
modSE0.ml <- lme(SE ~ 1,
              random=list(dummy=pdBlocked(list(pdIdent(~Composition-1), 
                                               pdIdent(~Block-1)))),
              data=NE_data,
              weights=varIdent(form=~ 1 | Days),
              method="ML", 
              control=list(msMaxIter=1000,
                             msMaxEval=1000))

anova(modSE0.ml, modSE.ml) #overall, the factor days has a statistically significant 
#effect on complementarity effect

summary(modSE)
#here we can see that CE is significantly greater at day 40 than at day 20, but CE 
#isn't significantly greater at day 60 than at day 20.

contrast(emmeans(modSE, specs="Days"), "pairwise") #the effect of days on SE is only 
#different at Day 40

par(mfrow=c(1,2))
plot(fitted(modSE), resid(modSE,type="pearson"))
qqnorm(resid(modSE,type="pearson"))
qqline(resid(modSE,type="pearson"))

meanSE<-as.vector(c(mean(NE_data[NE_data$Days==20, ]$SE), mean(NE_data[NE_data$Days==40, ]$SE), mean(NE_data[NE_data$Days==60, ]$SE)))

seSE<-as.vector(c(sd(NE_data[NE_data$Days==20, ]$SE)/sqrt(length(NE_data[NE_data$Days==20, ]$SE)), sd(NE_data[NE_data$Days==40, ]$SE)/sqrt(length(NE_data[NE_data$Days==40, ]$SE)), sd(NE_data[NE_data$Days==60, ]$SE)/sqrt(length(NE_data[NE_data$Days==60, ]$SE))))
                                                                                                                                  
SE_meanse<-data.frame(
  days=c("20", "40", "60"),
  mean=meanSE,
  se=seSE,
  label=c("a", "b", "a"))
SE_meanse$ci<-SE_meanse$se*1.96
SE_meanse$stars<-SE_meanse$mean + SE_meanse$ci + 125

SEp<-ggplot(SE_meanse) +
  geom_bar(aes(x=days, y=mean), colour="black", fill="#2D3184", stat="identity", alpha=0.7) +
  #geom_errorbar(aes(x=days, ymin=mean-se, ymax=mean+se), width=0.4, colour="red", alpha=0.9, size=1.3) +
  geom_errorbar(aes(x=days, ymin=mean-ci, ymax=mean+ci), width=0.4, colour="black", alpha=0.9, size=1.3) +
  geom_text(aes(x=days, y=mean+ci+100, label=label,), position=position_dodge(0.9), vjust=1, size=7) +
  theme_classic() + 
  theme(text=element_text(size=15), plot.title=element_text(vjust=2), panel.grid.major.y = element_line( size=.1, color="black" )) +
  labs(x="Days of growth", y="SE", title="Selection effect", pch=8) +
  scale_y_continuous(limits=c(-400, 600), breaks=seq(-400, 600, by=200)) + 
  annotate(
    "text",
    x = c(1, 2, 3),  
    y = c(161.9431, 326.8847, 146.8014), 
    label = c("*", "****", "n.s."), 
    size = 5,  
    color = "black"  
  )
```
### Two-tailed t-tests.
Here we want to see if the global NE, CE and SE is different from 0 for each level of the factor Days.
```{r}
NE_ttests<-NE_data %>%
  group_by(Days) %>%
  t_test(NE ~ 0) %>%
  add_significance()
NE_ttests

CE_ttests<-NE_data %>%
  group_by(Days) %>%
  t_test(CE ~ 0) %>%
  add_significance()
CE_ttests

SE_ttests<-NE_data %>%
  group_by(Days) %>%
  t_test(SE ~ 0) %>%
  add_significance()
SE_ttests
```
###Final plots BEF through time
```{r}
BEFplots<-ggarrange(NEp, CEp, SEp, nrow=1, ncol=3, common.legend = TRUE) 
BEFplots
```
