---
title: "2. Functional Diversity"
author: "meganedeziel"
date: "2024-02-06"
output:
  pdf_document: default
  html_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, message=FALSE, warning=FALSE)
```

This code calculates Fundamental and Realized Functional diversity 
considering leaf area and runs models 

```{r, message=FALSE, warning=FALSE}
library(here)
library(tidyverse)
library(FD)
library(nlme)
library(MuMIn)
library(ggpubr)
```

### Data importation
```{r}
#Trait data
Biom_data<-read.csv(here("data", "Relativebiomass_data.csv"), sep=";")

#Data on Net effect, CE, SE 
NE_data<-read.csv(here("data", "NE_data.csv"), sep=";")

#transform Composition, Block, Days as factor for further analyses
NE_data<-NE_data[NE_data$Days!="0", -1]
NE_data<-NE_data %>%
  mutate(across(c("Composition", "Block", "Days"), 
                as.factor)) 
NE_data$dummy<-factor(1)
NE60<-NE_data[NE_data$Days==60,]
```

### Graphs : Difference in leaf area of target species between Day 0 and Day 60 in every mixture
```{r}
Lm_la_data<-Biom_data[Biom_data$Species=="Lm" & Biom_data$Days %in% c(0, 60), ] %>%
  dplyr::select(., Block, Mixture, Days, Mean_LA) %>%
  pivot_wider(., names_from=Days, values_from = Mean_LA)
Lm_la_data$difference<-Lm_la_data$'60'-Lm_la_data$'0'
Lm_la_data$otherspecies<-rep(c("Lm", "Sp", "Wc", "Lt", "SpWc", "SpLt", "WcLt", "SpWcLt"), 3)
Lm_la_data$otherspecies<-fct_relevel(Lm_la_data$otherspecies, c("Lm", "Sp", "Wc", "Lt", "SpWc", "SpLt", "WcLt", "SpWcLt"))

Lm_boxplot<-
  ggplot(Lm_la_data, aes(x = otherspecies, y = difference)) + 
  geom_boxplot(fill="#2D3184", alpha=0.7) +
  geom_point() +
  ggtitle("Lm") +
  theme_classic() +
  theme(axis.title.x=element_blank(),
        axis.title.y=element_blank(),
        text=element_text(size=15), plot.title=element_text(vjust=2), panel.grid.major.y = element_line(linewidth=.1, color="black" ))

Sp_la_data<-Biom_data[Biom_data$Species=="Sp" & Biom_data$Days %in% c(0, 60), ] %>%
  dplyr::select(., Block, Mixture, Days, Mean_LA) %>%
  pivot_wider(., names_from=Days, values_from = Mean_LA)
Sp_la_data$difference<-Sp_la_data$'60'-Sp_la_data$'0'
Sp_la_data$otherspecies<-rep(c("Sp", "Lm", "Wc", "Lt", "LmWc", "LmLt", "WcLt", "LmWcLt"), 3)
Sp_la_data$otherspecies<-fct_relevel(Sp_la_data$otherspecies, c("Sp", "Lm", "Wc", "Lt", "LmWc", "LmLt", "WcLt", "LmWcLt"))

Sp_boxplot<-
  ggplot(Sp_la_data, aes(x = otherspecies, y = difference)) + 
  geom_boxplot(fill="#2D3184", alpha=0.7) +
  geom_point() +
  ggtitle("Sp") +
  theme_classic() +
  theme(axis.title.x=element_blank(),
        axis.title.y=element_blank(),
        text=element_text(size=15), plot.title=element_text(vjust=2), panel.grid.major.y = element_line(linewidth=.1, color="black" ))

Wc_la_data<-Biom_data[Biom_data$Species=="Wc" & Biom_data$Days %in% c(0, 60), ] %>%
  dplyr::select(., Block, Mixture, Days, Mean_LA) %>%
  pivot_wider(., names_from=Days, values_from = Mean_LA)
Wc_la_data$difference<-Wc_la_data$'60'-Wc_la_data$'0'
Wc_la_data$otherspecies<-rep(c("Wc", "Lm", "Sp", "Lt", "LmSp", "LmLt", "SpLt", "LmSpLt"), 3)
Wc_la_data$otherspecies<-fct_relevel(Wc_la_data$otherspecies, c("Wc", "Lm", "Sp", "Lt", "LmSp", "LmLt", "SpLt", "LmSpLt"))

Wc_boxplot<-
  ggplot(Wc_la_data, aes(x = otherspecies, y = difference)) + 
  geom_boxplot(fill="#2D3184", alpha=0.7) +
  geom_point() +
  ggtitle("Wc") +
  theme_classic() +
  theme(axis.title.x=element_blank(),
        axis.title.y=element_blank(),
        text=element_text(size=15), plot.title=element_text(vjust=2), panel.grid.major.y = element_line(linewidth=.1, color="black" ))

Lt_la_data<-Biom_data[Biom_data$Species=="Lt" & Biom_data$Days %in% c(0, 60), ] %>%
  dplyr::select(., Block, Mixture, Days, Mean_LA) %>%
  pivot_wider(., names_from=Days, values_from = Mean_LA)
Lt_la_data$difference<-Lt_la_data$'60'-Lt_la_data$'0'
Lt_la_data$otherspecies<-rep(c("Lt", "Lm", "Sp", "Wc", "LmSp", "LmWc", "SpWc", "LmSpWc"), 3)
Lt_la_data$otherspecies<-fct_relevel(Lt_la_data$otherspecies, c("Lt", "Lm", "Sp", "Wc", "LmSp", "LmWc", "SpWc", "LmSpWc"))


Lt_boxplot<-
  ggplot(Lt_la_data, aes(x = otherspecies, y = difference)) + 
  geom_boxplot(fill="#2D3184", alpha=0.7) + #aes(middle = mean(difference))
  geom_point() +
  ggtitle("Lt") +
  theme_classic() +
  theme(axis.title.x=element_blank(),
        axis.title.y=element_blank(),
        text=element_text(size=15), plot.title=element_text(vjust=2), panel.grid.major.y = element_line(linewidth=.1, color="black" ))


difference<-ggarrange(Lm_boxplot, Sp_boxplot, Wc_boxplot, Lt_boxplot)
annotate_figure(difference,
                left=text_grob(expression("Difference in leaf area (mm"^2*")"), rot=90, size=15))
```

### Functional diversity calculations
#### Abundance matrices for further calculations
```{r}
Biom_data$MixBlock<-paste(Biom_data$Mixture, Biom_data$Block, sep="_")
Biom_data$MixBlockSp<-paste(Biom_data$Species, Biom_data$Mixture, Biom_data$Block, sep="_")

a_matrix<-subset(Biom_data, Biom_data$Days==0 & !(Biom_data$Mixture %in% c("Lm", "Sp", "Wc", "Lt"))) %>%
  dplyr::select(., MixBlock, Species, Initial_relbiomass) %>%
  unique() %>%
  pivot_wider(names_from=Species, values_from=Initial_relbiomass) %>% #NAs will be replaced by 0
  column_to_rownames(var="MixBlock") %>%
  as.matrix()
a_matrix_block1<-a_matrix[1:11,]
a_matrix_block2<-a_matrix[12:22,]
a_matrix_block3<-a_matrix[23:33,]

a_matrix_poly<-subset(Biom_data, Biom_data$Days==0 & !(Biom_data$Mixture %in% c("Lm", "Sp", "Wc", "Lt"))) %>%
  dplyr::select(., MixBlock, MixBlockSp, Initial_relbiomass) %>%
  unique() %>%
  pivot_wider(names_from=MixBlockSp, values_from=Initial_relbiomass) %>%
  column_to_rownames(var="MixBlock") %>%
  as.matrix()
a_matrix_poly_block1<-a_matrix_poly[1:11, 1:28]
a_matrix_poly_block2<-a_matrix_poly[12:22, 29:56]
a_matrix_poly_block3<-a_matrix_poly[23:33, 57:84]
```

#### FFD
```{r}
#FFD
#mean value of leaf area when Days==0; one value per species
d_matrix_ffd<-Biom_data[Biom_data$Days==0 & Biom_data$Mixture %in% unique(Biom_data$Species), ] %>%
  group_by(Mixture) %>%
  mutate(mean3_LA=mean(Mean_LA)) %>%
  dplyr::select(., Mixture, mean3_LA) %>%
  unique() %>%
  column_to_rownames(var="Mixture")
colnames(d_matrix_ffd)<-"LA"

ffd<-as.data.frame(fdisp(dist(d_matrix_ffd), a_matrix)$FDis)
colnames(ffd)<-"FFD"
```

#### FFD60
```{r}
#FFD60 ; here we use one value of leaf area per species (leaf area = mean of leaf area computed from the values of each of the 3 blocks at 60 days)
d_matrix_ffd60<-Biom_data[Biom_data$Days==60 & Biom_data$Mixture %in% unique(Biom_data$Species), ] %>%
  group_by(Mixture) %>%
  mutate(mean3_LA=mean(Mean_LA)) %>%
  dplyr::select(., Mixture, mean3_LA) %>%
  unique() %>%
  column_to_rownames(var="Mixture")
colnames(d_matrix_ffd60)<-"LA"

ffd60<-as.data.frame(fdisp(dist(d_matrix_ffd60), a_matrix)$FDis)
colnames(ffd60)<-"FFD60"

#FFD60 (PER BLOCK) ; here we consider the different blocks (leaf area = raw value at 60 days for each monoculture)
#block 1
d_matrix_ffd60_Bl1<-Biom_data[Biom_data$Days==60 & Biom_data$Mixture %in% unique(Biom_data$Species) & Biom_data$Block==1, ] %>%
  dplyr::select(., Mixture, Mean_LA) %>%
  unique()
row.names(d_matrix_ffd60_Bl1)<-as.vector(d_matrix_ffd60_Bl1$Mixture) 
d_matrix_ffd60_Bl1<-d_matrix_ffd60_Bl1 %>% dplyr::select(., Mean_LA)
colnames(d_matrix_ffd60_Bl1)<-"LA"

ffd60_Bl1<-as.data.frame(fdisp(dist(d_matrix_ffd60_Bl1), a_matrix_block1)$FDis)
colnames(ffd60_Bl1)<-"FFD60_perblock"

#block 2
d_matrix_ffd60_Bl2<-Biom_data[Biom_data$Days==60 & Biom_data$Mixture %in% unique(Biom_data$Species) & Biom_data$Block==2, ] %>%
  dplyr::select(., Mixture, Mean_LA) %>%
  unique()
row.names(d_matrix_ffd60_Bl2)<-as.vector(d_matrix_ffd60_Bl2$Mixture) 
d_matrix_ffd60_Bl2<-d_matrix_ffd60_Bl2 %>% dplyr::select(., Mean_LA)
colnames(d_matrix_ffd60_Bl2)<-"LA"

ffd60_Bl2<-as.data.frame(fdisp(dist(d_matrix_ffd60_Bl2), a_matrix_block2)$FDis)
colnames(ffd60_Bl2)<-"FFD60_perblock"

#block 3
d_matrix_ffd60_Bl3<-Biom_data[Biom_data$Days==60 & Biom_data$Mixture %in% unique(Biom_data$Species) & Biom_data$Block==3, ] %>%
  dplyr::select(., Mixture, Mean_LA) %>%
  unique()
row.names(d_matrix_ffd60_Bl3)<-as.vector(d_matrix_ffd60_Bl3$Mixture) 
d_matrix_ffd60_Bl3<-d_matrix_ffd60_Bl3 %>% dplyr::select(., Mean_LA)
colnames(d_matrix_ffd60_Bl3)<-"LA"

ffd60_Bl3<-as.data.frame(fdisp(dist(d_matrix_ffd60_Bl3), a_matrix_block3)$FDis)
colnames(ffd60_Bl3)<-"FFD60_perblock"

#put together
ffd60_perblock<-rbind(ffd60_Bl1, ffd60_Bl2, ffd60_Bl3)
```

### RFD
```{r}
# RFD ; here we consider each culture seperately
# block 1 
d_matrix_rfd_Bl1<-Biom_data[Biom_data$Days==60 & !(Biom_data$Mixture %in% c("Lm", "Sp", "Wc", "Lt")) & Biom_data$Block==1, ] %>%
  dplyr::select(., MixBlockSp, Mean_LA) %>%
  unique()
row.names(d_matrix_rfd_Bl1)<-as.vector(d_matrix_rfd_Bl1$MixBlockSp) 
d_matrix_rfd_Bl1<-d_matrix_rfd_Bl1 %>% dplyr::select(., Mean_LA)
colnames(d_matrix_rfd_Bl1)<-"LA"

rfd_Bl1<-as.data.frame(fdisp(dist(d_matrix_rfd_Bl1), a_matrix_poly_block1)$FDis)
colnames(rfd_Bl1)<-"RFD"

#block 2
d_matrix_rfd_Bl2<-Biom_data[Biom_data$Days==60 & !(Biom_data$Mixture %in% c("Lm", "Sp", "Wc", "Lt")) & Biom_data$Block==2, ] %>%
  dplyr::select(., MixBlockSp, Mean_LA) %>%
  unique()
row.names(d_matrix_rfd_Bl2)<-as.vector(d_matrix_rfd_Bl2$MixBlockSp) 
d_matrix_rfd_Bl2<-d_matrix_rfd_Bl2 %>% dplyr::select(., Mean_LA)
colnames(d_matrix_rfd_Bl2)<-"LA"

rfd_Bl2<-as.data.frame(fdisp(dist(d_matrix_rfd_Bl2), a_matrix_poly_block2)$FDis)
colnames(rfd_Bl2)<-"RFD"

#block 3
d_matrix_rfd_Bl3<-Biom_data[Biom_data$Days==60 & !(Biom_data$Mixture %in% c("Lm", "Sp", "Wc", "Lt")) & Biom_data$Block==3, ] %>%
  dplyr::select(., MixBlockSp, Mean_LA) %>%
  unique()
row.names(d_matrix_rfd_Bl3)<-as.vector(d_matrix_rfd_Bl3$MixBlockSp) 
d_matrix_rfd_Bl3<-d_matrix_rfd_Bl3 %>% dplyr::select(., Mean_LA)
colnames(d_matrix_rfd_Bl3)<-"LA"

rfd_Bl3<-as.data.frame(fdisp(dist(d_matrix_rfd_Bl3), a_matrix_poly_block3)$FDis)
colnames(rfd_Bl3)<-"RFD"

#put together
rfd<-rbind(rfd_Bl1, rfd_Bl2, rfd_Bl3)
```

#### Combine all FD indices
```{r}
fd<-cbind(ffd, ffd60, ffd60_perblock, rfd)
```

### Models: Effect of functional diversity indices on NE and CE
```{r}
fd_NE<-cbind(fd, NE60)

#model NE ~ FFD
modNEFFD<-lme(NE ~ FFD, random=list(dummy = pdBlocked(list(pdIdent(~Composition-1), pdIdent(~Block-1)))), data=fd_NE)
modNEFFD
summary(modNEFFD)
r.squaredGLMM(modNEFFD)
plot(NE ~ FFD, data = fd_NE) 
abline(modNEFFD)

par(mfrow=c(1,2))
plot(fitted(modNEFFD), resid(modNEFFD,type="pearson"))
qqnorm(resid(modNEFFD,type="pearson"))
qqline(resid(modNEFFD,type="pearson"))

#alternative model without Composition as random factor
modaltNEFFD<-lme(NE ~ FFD, random=~1|Block, data=fd_NE)
modaltNEFFD
summary(modaltNEFFD)
r.squaredGLMM(modaltNEFFD)
par(mfrow=c(1,2))
plot(fitted(modaltNEFFD), resid(modaltNEFFD,type="pearson"))
qqnorm(resid(modaltNEFFD,type="pearson"))
qqline(resid(modaltNEFFD,type="pearson"))
```

```{r}
#model NE ~ FFD2
modNEFFD2<-lme(NE ~ FFD60_perblock, random=list(dummy = pdBlocked(list(pdIdent(~Composition-1), pdIdent(~Block-1)))), data=fd_NE)
modNEFFD2
summary(modNEFFD2)
r.squaredGLMM(modNEFFD2)
plot(NE ~ FFD60_perblock, data = fd_NE) 
abline(modNEFFD2)

par(mfrow=c(1,2))
plot(fitted(modNEFFD2), resid(modNEFFD2,type="pearson"))
qqnorm(resid(modNEFFD2,type="pearson"))
qqline(resid(modNEFFD2,type="pearson"))

#alternative model without Composition as random factor
modaltNEFFD2<-lme(NE ~ FFD60_perblock, random=~1|Block, data=fd_NE)
modaltNEFFD2
summary(modaltNEFFD2)
r.squaredGLMM(modaltNEFFD2)
par(mfrow=c(1,2))
plot(fitted(modaltNEFFD2), resid(modaltNEFFD2,type="pearson"))
qqnorm(resid(modaltNEFFD2,type="pearson"))
qqline(resid(modaltNEFFD2,type="pearson"))
```


```{r}
#model NE ~ RFD
modNERFD<-lme(NE ~ RFD, random=list(dummy = pdBlocked(list(pdIdent(~Composition-1), pdIdent(~Block-1)))), data=fd_NE)
modNERFD
summary(modNERFD)
r.squaredGLMM(modNERFD)
plot(NE ~ FFD60, data = fd_NE) 
abline(modNEFFD2_m)

par(mfrow=c(1,2))
plot(fitted(modNERFD), resid(modNERFD,type="pearson"))
qqnorm(resid(modNERFD,type="pearson"))
qqline(resid(modNERFD,type="pearson"))

#alternative model without Composition as random factor
modaltNERFD<-lme(NE ~ RFD, random=~1|Block, data=fd_NE)
modaltNERFD
summary(modaltNERFD)
r.squaredGLMM(modaltNERFD)

par(mfrow=c(1,2))
plot(fitted(modaltNERFD), resid(modaltNERFD,type="pearson"))
qqnorm(resid(modaltNERFD,type="pearson"))
qqline(resid(modaltNERFD,type="pearson"))
```

```{r}
#model CE ~ FFD
modCEFFD<-lme (CE ~ FFD, random=list(dummy = pdBlocked(list(pdIdent(~Composition-1), pdIdent(~Block-1)))), data=fd_NE)
modCEFFD
summary(modCEFFD)
r.squaredGLMM(modCEFFD)
plot(CE ~ FFD, data = fd_NE) 
abline(modCEFFD)

par(mfrow=c(1,2))
plot(fitted(modCEFFD), resid(modCEFFD,type="pearson"))
qqnorm(resid(modCEFFD,type="pearson"))
qqline(resid(modCEFFD,type="pearson"))

#alternative model without Composition as random factor
modaltCEFFD<-lme(CE ~ FFD, random=~1|Block, data=fd_NE)
modaltCEFFD
summary(modaltCEFFD)
r.squaredGLMM(modaltCEFFD)
par(mfrow=c(1,2))
plot(fitted(modaltCEFFD), resid(modaltCEFFD,type="pearson"))
qqnorm(resid(modaltCEFFD,type="pearson"))
qqline(resid(modaltCEFFD,type="pearson"))
```

```{r}
#model CE ~ FFD2
modCEFFD2<-lme (CE ~ FFD60_perblock, random=list(dummy = pdBlocked(list(pdIdent(~Composition-1), pdIdent(~Block-1)))), data=fd_NE)
modCEFFD2
summary(modCEFFD2)
r.squaredGLMM(modCEFFD2)
plot(CE ~ FFD60_perblock, data = fd_NE) 
abline(modCEFFD2)

par(mfrow=c(1,2))
plot(fitted(modCEFFD2), resid(modCEFFD2,type="pearson"))
qqnorm(resid(modCEFFD2,type="pearson"))
qqline(resid(modCEFFD2,type="pearson"))

#alternative model without Composition as random factor
modaltCEFFD2<-lme(CE ~ FFD60_perblock, random=~1|Block, data=fd_NE)
modaltCEFFD2
summary(modaltCEFFD2)
r.squaredGLMM(modaltCEFFD2)
par(mfrow=c(1,2))
plot(fitted(modaltCEFFD2), resid(modaltCEFFD2,type="pearson"))
qqnorm(resid(modaltCEFFD2,type="pearson"))
qqline(resid(modaltCEFFD2,type="pearson"))
```

```{r}
#model CE ~ RFD
modCERFD<-lme (CE ~ RFD, random=list(dummy = pdBlocked(list(pdIdent(~Composition-1), pdIdent(~Block-1)))), data=fd_NE)
modCERFD<-lme (CE ~ RFD, random=list(dummy = pdBlocked(list(pdIdent(~Composition-1), pdIdent(~Block-1)))), data=fd_NE)
modCERFD
summary(modCERFD)
r.squaredGLMM(modCERFD)
plot(CE ~ RFD, data = fd_NE) 
abline(modCERFD)

par(mfrow=c(1,2))
plot(fitted(modCERFD), resid(modCERFD,type="pearson"))
qqnorm(resid(modCERFD,type="pearson"))
qqline(resid(modCERFD,type="pearson"))

#alternative model without Composition as random factor
modaltCERFD<-lme(CE ~ RFD, random=~1|Block, data=fd_NE)
modaltCERFD
summary(modaltCERFD)
r.squaredGLMM(modaltCERFD)
par(mfrow=c(1,2))
plot(fitted(modaltCERFD), resid(modaltCERFD,type="pearson"))
qqnorm(resid(modaltCERFD,type="pearson"))
qqline(resid(modaltCERFD,type="pearson"))
```