---
title: "4. Two-tailed t-tests"
author: "meganedeziel"
date: "2024-01-25"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

This code aims to test the hypothesis that Net effect is different (less than OR greater than) from 0 in each of the mixtures at different levels of the factor Days.

### Package loading
```{r, message=FALSE}
library(here)
library(tidyverse)
library(rstatix)
```

### Data importation 
```{r}
NE_data<-read.csv(here("data", "NE_data.csv"), sep=";")
```

Null hypothesis: There is no difference between the sample mean and the hypothesized population mean (0)
Alternative hypothesis: Diversity has a significant effect on net effect (NE != 0)
### Tests on NE
```{r}
NE60<-NE_data[NE_data$Days=="60",] %>%
  group_by(Composition) %>%
  t_test(NE ~ 0) %>%
  add_significance()
NE60
meanNE60<-NE_data[NE_data$Days=="60",] %>%
  group_by(Composition) %>%
  mutate(mean_60=mean(NE)) %>%
  unique()

NE40<-NE_data[NE_data$Days=="40",] %>%
  group_by(Composition) %>%
  t_test(NE ~ 0) %>%
  add_significance()
NE40
meanNE40<-NE_data[NE_data$Days=="40",] %>%
  group_by(Composition) %>%
  mutate(mean_40=mean(NE)) %>%
  unique()

NE20<-NE_data[NE_data$Days=="20",] %>%
  group_by(Composition) %>%
  t_test(NE ~ 0) %>%
  add_significance()
NE20
meanNE20<-NE_data[NE_data$Days=="20",] %>%
  group_by(Composition) %>%
  mutate(mean_20=mean(NE)) %>%
  unique()
```

### Tests on CE
```{r}
CE60<-NE_data[NE_data$Days=="60",] %>%
  group_by(Composition) %>%
  t_test(CE ~ 0) %>%
  add_significance()
CE60
meanCE60<-NE_data[NE_data$Days=="60",] %>%
  group_by(Composition) %>%
  mutate(mean_60=mean(CE)) %>%
  unique()

CE40<-NE_data[NE_data$Days=="40",] %>%
  group_by(Composition) %>%
  t_test(CE ~ 0) %>%
  add_significance()
CE40
meanCE40<-NE_data[NE_data$Days=="40",] %>%
  group_by(Composition) %>%
  mutate(mean_40=mean(CE)) %>%
  unique()

CE20<-NE_data[NE_data$Days=="20",] %>%
  group_by(Composition) %>%
  t_test(CE ~ 0) %>%
  add_significance()
CE20
meanCE20<-NE_data[NE_data$Days=="20",] %>%
  group_by(Composition) %>%
  mutate(mean_20=mean(CE)) %>%
  unique()
```

### Tests on SE
```{r}
SE60<-NE_data[NE_data$Days=="60",] %>%
  group_by(Composition) %>%
  t_test(SE ~ 0) %>%
  add_significance()
SE60
meanSE60<-NE_data[NE_data$Days=="60",] %>%
  group_by(Composition) %>%
  mutate(mean_60=mean(CE)) %>%
  unique()

SE40<-NE_data[NE_data$Days=="40",] %>%
  group_by(Composition) %>%
  t_test(SE ~ 0) %>%
  add_significance()
SE40
meanSE40<-NE_data[NE_data$Days=="40",] %>%
  group_by(Composition) %>%
  mutate(mean_40=mean(SE)) %>%
  unique()

SE20<-NE_data[NE_data$Days=="20",] %>%
  group_by(Composition) %>%
  t_test(SE ~ 0) %>%
  add_significance()
SE20
meanSE20<-NE_data[NE_data$Days=="20",] %>%
  group_by(Composition) %>%
  mutate(mean_20=mean(SE)) %>%
  unique()
```



