---
title: "Average_biomass_per_species"
author: "meganedeziel"
date: "2023-10-21"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```
```{r}
library("here")
library("tidyverse")
This script will produce Figure 2: The relative contribution of species to the average biomass on day 60 of the experiment
```{r}
Biom_data<-read.csv(here("data", "Relativebiomass_data.csv"), sep=";")
#info on data
#individual mass : in mg
#mean LA = mean leaf area : in mg
#mean SLA = mean specific leaf area : in mm2/mg
#initial relbiomass : in mg
#final relbiomass: in mg

Biom_data<-Biom_data %>%
  mutate(across(c("Culture", "Block", "Mixture", "Species"), 
                as.factor)) %>%
  group_by(Mixture)
  mutate(Biomass_tot_per_mix<-)
```

```{r}
Plot_data<-Biom_data[Biom_data$Days==60,] %>%
  group_by(Mixture, Block) %>% #sum per Mixture per Block
  mutate(sum_finbiomass=sum(Final_relbiomass)) %>% #i now have the total biomass per culture
  ungroup() %>%
  group_by(Mixture) %>% #to calculate mean per mixture, considering the 3 blocks
  
  group_by(Mixture) %>%
  mutate(mean_finbiomass_per_mix=mean(Final_relbiomass))
ggplot(Biom_data, aes(fill=Species, y=Final_relbiomass, x=Mixture)) +
    geom_bar(position="stack", stat="identity")

```

